`r glue::glue('Initial inferCNV results for {params$library_id}')`

### 2025-04-15 analyses/infercnv-consensus-cell-type/template-notebooks/ewings-infercnv-results.Rmd
Add a plot to directly compare references

* `all-immune`: All consensus immune cells across `SCPCP000015` libraries, _excluding_ any which were also identified as tumor in the `cell-type-ewings` analysis module
* `subset-immnune`: A subset of `all-immune`, considering only macrophage and T-cell cell types

@@

* We used the `i3` HMM which has improved run times relative to their `i6` HMM
* We used the default `analysis_mode = "subclusters"`, which allows the HMM to call CNVs at the level of clusters rather than whole samples or individual cells
* We used the following clustering parameters:
  * Leiden clustering run on PCA (not the expression matrix)
  * `k=20` nearest neighbors with a `resolution` parameter of 1 and using the `modularity` objective function

@@

# Set default ggplot theme, customized for UMAPs
theme_set(theme_bw())

@@

@@

```{r}
# First, construct data frame of total `has_cnv` counts, with other variables needed for plots
has_cnv_df <- infercnv_df |>
  dplyr::select(
    barcodes,
    normal_reference,
    subcluster,
    starts_with("has_cnv_")
  ) |>
  tidyr::pivot_longer(
    starts_with("has_cnv_"),
    names_to = "chr",
    values_to = "cnv"
  ) |>
  # sum all values for each cell, but separately for each run of inferCNV (aka each normal reference)
  dplyr::group_by(barcodes, cell_id, normal_reference) |>
  dplyr::summarize(total_cnv_per_cell = sum(cnv)) |>
  # add indicator columns for whether the cell is reference or query
  dplyr::mutate(
    immune_reference_cell = barcodes %in% immune_ref_barcodes,
    subset_reference_cell = barcodes %in% subset_ref_barcodes
  ) |>
  # join in other data
  dplyr::left_join(celltype_df) |>
  dplyr::left_join(umap_df)
```

@@

First, as a guide, let's look at the top 10 `cell-type-ewings` cell types.

@@

>>>>

* `all-immune`: All consensus immune cells across `SCPCP000015` libraries, _excluding_ any which were also identified as tumor in the `cell-type-ewings` analysis module
* `subset-immune`: A subset of `all-immune`, considering only macrophage and T-cell cell types

@@

* We used the `i3` HMM which has improved run times relative to the `i6` HMM.
* We used the default `analysis_mode = "subclusters"`, which allows the HMM to call CNVs at the level of clusters rather than whole samples or individual cells
* We used the following clustering parameters:
  * Leiden clustering run on PCA (not the expression matrix)
  * `k=20` nearest neighbors with a `resolution` parameter of 1 and using the `modularity` objective function

@@

theme_set(theme_bw())

@@

To help contextualize these, here is the table of known CNVs in Ewing sarcoma we'll use for validation:

@@

```{r}
# First, construct data frame of total `has_cnv` counts, with other variables needed for plots
has_cnv_df <- infercnv_df |>
  dplyr::select(
    barcodes,
    normal_reference,
    subcluster,
    starts_with("has_cnv_")
  ) |>
  tidyr::pivot_longer(
    starts_with("has_cnv_"),
    names_to = "chr",
    values_to = "cnv"
  ) |>
  # sum all values for each cell, but separately for each run of inferCNV (aka each normal reference)
  dplyr::group_by(barcodes, normal_reference) |>
  dplyr::summarize(total_cnv_per_cell = sum(cnv)) |>
  # add indicator columns for whether the cell is reference or query
  dplyr::mutate(
    immune_reference_cell = barcodes %in% immune_ref_barcodes,
    subset_reference_cell = barcodes %in% subset_ref_barcodes
  ) |>
  # join in other data
  dplyr::left_join(celltype_df) |>
  dplyr::left_join(umap_df)
```

@@

First, as a guide to interpret the UMAP, let's color by the top 10 `cell-type-ewings` cell types.

@@

## Compare references directly

We can also look at CNV predictions directly between references using a point-density scatterplot, with the Spearman correlation shown:

```{r}
compare_refs <- has_cnv_df |>
  # bring in library_id so we can only include those cells.
  dplyr::inner_join(
    infercnv_df |>
      dplyr::select(barcodes, library_id) |>
      unique()
  ) |>
  dplyr::filter(library_id == params$library_id) |>
  dplyr::select(
    barcodes, normal_reference, total_cnv_per_cell
  ) |>
  tidyr::pivot_wider(
    names_from = normal_reference,
    values_from = total_cnv_per_cell
  ) |>
  # color points this way rather than binning to ensure small bins aren't left out
  dplyr::add_count(`all-immune`, `subset-immune`)

# Determine max limit for axes
max_axis <- max(has_cnv_df$total_cnv_per_cell)
axis_breaks <- 1:max_axis

# Calculate in spearman correlation coefficient to include in plot
rho <- cor(compare_refs$`all-immune`, compare_refs$`subset-immune`, method = "spearman") |>
  round(2)

ggplot(compare_refs) +
  aes(
    x = `all-immune`,
    y = `subset-immune`,
    color = n
  ) +
  geom_point(size = 5) +
  geom_abline(color = "red") +
  scale_color_viridis_c() +
  coord_equal() +
  scale_x_continuous(limits = c(0, max_axis), breaks = 1:max_axis) +
  scale_y_continuous(limits = c(0, max_axis), breaks = 1:max_axis) +
  labs(
    title = "Per-cell CNVs between normal references",
    subtitle = glue::glue("rho = {rho}"),
    color = "Number of cells"
  ) +
  theme(legend.position = "bottom")
```
